---
title: "MIL_explore"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

read 50 year data, change date column data type <- not working
```{r}

library(tidyverse)

data <- read_csv("MIL_50years_R.csv")

colnames(data) <- c("id", "duration", "sensor_number", "sensor_type", "year", "month", "day", "date", "obs_date", "value", "flag", "units")


as.Date(data$date, format = "%m-%d-%y")

as.Date(data$obs_date, format = "%Y-%m-%d")



inflow <- data %>% 
  filter(sensor_type == "INFLOW")
as.Date.factor(inflow$date, format = "%m-%d-%y")


outflow <- data %>% 
  filter(sensor_type == "OUTFLOW")


release<- data %>% 
  filter(sensor_type == "REL SCH")


storage<- data %>% 
  filter(sensor_type == "STORAGE") %>% 
  select("id", "obs_date", "value", "units") %>% 
  na.omit(value) %>% 
  mutate(percentage = (value/520528)*100)


storage$percentage<-round(storage$percentage, 2)




```



test run?
```{r}

plot <- ggplot(data = inflow, aes(x = obs_date, y = value))+
  geom_line()+
  geom_smooth()+
  theme_minimal()+
  scale_y_continuous(limits = c(0,50000))

plot


```




```{r}


plot_storage <- ggplot(data = storage, aes(x = obs_date, y = value))+
  geom_line()+
  geom_smooth()+
  theme_minimal()+
  scale_y_continuous(limits = c(0,600000))

plot_storage


plot_percent <- ggplot(data = storage, aes(x = obs_date, y = percentage))+
  geom_area()+
  theme_minimal()+
  scale_y_continuous(limits = c(0,120))

plot_percent


```



```{r}

plot_outflow <- ggplot(data = outflow, aes(x = obs_date, y = value))+
  geom_line()+
  geom_smooth()+
  theme_minimal()+
  scale_y_continuous(limits = c(0,35000))


plot_outflow


```



```{r}


plot_release <- ggplot(data = release, aes(x = obs_date, y = value))+
  geom_point()+
  theme_minimal()+
  scale_y_continuous(limits = c(0,35000))


plot_release





```


try with last 10 years of data for simpler visualization
```{r}

storage10 <- subset(storage, obs_date > as.Date ("2009-1-1"))
outflow10 <- subset(outflow, obs_date > as.Date ("2009-1-1"))



plot_storage10 <- ggplot(data = storage10, aes(x = obs_date, y = value))+
  geom_line()+
  geom_smooth()+
  theme_minimal()+
  scale_y_continuous(limits = c(0,600000))

plot_storage10


plot_percent10 <- ggplot()+
  geom_area(data = storage10, aes(x = obs_date, y = value))+
  geom_line(data = outflow10, aes(x = obs_date, y = value))+
  theme_minimal()+
  scale_y_continuous(
  sec.axis = sec_axis(~., name = "outflows and releases (cfs)"),limits = c(0,20000))

plot_percent10




```




```{r}
storage2000 <- subset(storage, obs_date > as.Date ("2000-1-1"))
outflow2000 <- subset(outflow, obs_date > as.Date ("2000-1-1"))






```










